---
title: "Cluster by Enterotype"
output: 
  html_notebook:
    codes: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(tibble)

# pink/blue/
grp <- c("ET_B", "ET_P")
grp.col <- c("#FF6000", "#0080C0")

stg <- c("Before", "After")
stg.col <- c("#6288BA", "#C25D97")
```

### load data
```{r}
phen <- read.csv("../datset/phenotype.csv")
gsf <- read.table("../datset/profile/Genus.profile")
```

### genus profile in before for  enterotype classification 
[Enterotype reference](http://enterotypes.org/) attains Enterotype result
```{r}
before.phe <- phen %>% filter(Stage=="After")
before.gsf <- gsf[, colnames(gsf)%in%before.phe$SampleID]
colnames(before.gsf) <- before.phe$ID
write.table(before.gsf, "after.gsf.txt", quote = F, sep = "\t")

# enterotype by network reference
enterotype <- read.table("Enterotype_classification_network.txt",
                         header = T)
group <- inner_join(before.phe %>% select(ID),
                    enterotype %>% rownames_to_column("ID") %>%
                      select(ID, ET),
                    by="ID") %>% 
  dplyr::rename(Group=ET)

phe.grp <- inner_join(group %>% select(ID, Group), 
                      phen, by = "ID") %>%
  mutate(Group = factor(Group, levels = grp),
         Stage = factor(Stage, levels = stg))

```

### PERMANOVA 

validate that the enterotype is significantly differet
```{r}
idx <- c("SampleID", "ID", "Group", "Gender", "BMI", "Weight", "Age")
before.grp <- phe.grp %>% filter(Stage=="Before") %>% 
          select(one_of(idx))

PermProf <- function(x, y) {
  sid <- intersect(as.character(x$SampleID), colnames(y)) 
  phe <- x %>% filter(SampleID %in% sid) 
  prf <-  y %>% select(as.character(phe$SampleID)) %>%
          t() %>% data.frame()
  
  per <- apply(phe %>% select(-one_of("SampleID")), 2, function(a, pf){
    dat <- data.frame(value = a, pf)
      datphe <- dat$value
      if (length(datphe) == 0 | unique(datphe) == 1) {
        res <- data.frame(length(datphe), rep(NA, 6))
        next
      }
      
      if (length(unique(datphe)) < 6) {
        datphe <- as.factor(datphe)
      } 
      datprf <- dat[, -1, F] 
      library(vegan)
      dis <- vegdist(datprf, method = "bray")
      
      set.seed(123)
      ad <- adonis(dis ~ datphe, permutations = 1000)
      tmp <- as.data.frame(ad$aov.tab) %>% slice(1)
      res <- c(length(datphe), as.numeric(tmp[, c(1:6)]))
      return(res)
    }, prf) %>% t() %>% data.frame()
    
  colnames(per) <- c("SumsOfSample", "Df", "SumsOfSqs", 
                       "MeanSqs", "F.Model", "R2", "Pr(>F)")
  
  per$FDR <- p.adjust(per$`Pr(>F)`, method = "BH")
  
  return(per)
}

before.per <- PermProf(before.grp, gsf)
DT::datatable(before.per)
```


### Function 
```{r}
# subject information
CohortInform <- function(x){
  dat <- x %>% filter(Stage%in%stg[1]) %>% 
    select(ID, Age, Sex, Weight, BMI)
  name <- "cohort\nMean+/-SD"
  res <- dat %>% 
      summarise(
        `Number of subjects (n)` = length(ID),
        Sex = paste(paste0("F:", 
        round((length(Sex[Sex=="F"])/length(Sex)), 4)*100, "%"),
            paste0("M:", 
        round((length(Sex[Sex=="M"])/length(Sex)), 4)*100, "%")), 
        Age = paste0(round(mean(Age), 2), "+/-", round(sd(Age), 2)),
        Weight = paste0(round(mean(Weight), 2), "+/-", round(sd(Weight), 2)),
        BMI = paste0(round(mean(BMI), 2), "+/-", round(sd(BMI), 2))) %>% 
        t() %>% data.frame() %>% 
        setNames(name)
    
   return(res)
}

# Number of subject/Sex/Age/Weight/BMI
CohortNum <- function(x, tag){
 
  dat <- x %>% filter(Stage%in%stg[1]& Group%in%tag) %>% 
    select(ID, Age, Sex, Weight, BMI)
  name <- paste(tag, "cohort\nMean+/-SD")

  res <- dat %>% 
      summarise(
        `Number of subjects (n)` = length(ID),
        Sex = paste(paste0("F:", 
                round((length(Sex[Sex=="F"])/length(Sex)), 4)*100, "%"),
            paste0("M:", 
                round((length(Sex[Sex=="M"])/length(Sex)), 4)*100, "%")), 
        Age = paste0(round(mean(Age), 2), "+/-", round(sd(Age), 2)),
        Weight = paste0(round(mean(Weight), 2), "+/-", round(sd(Weight), 2)),
        BMI = paste0(round(mean(BMI), 2), "+/-", round(sd(BMI), 2))) %>% 
        t() %>% data.frame() %>% 
        setNames(name)
    
   return(res)
}

# pvalue of Sex/Age/Weight/BMI
CohortP <- function(x){
  dat <- x %>% filter(Stage%in%stg[1]) %>% 
    select(Age, Sex, Weight, BMI, Group)
  
  dat.cas <- dat %>% filter(Group%in%grp[1]) %>%
    select(-Group)
  dat.con <- dat %>% filter(Group%in%grp[2]) %>%
    select(-Group)
  
  Sex <- as.table(cbind(table(dat.cas$Sex), table(dat.con$Sex)))
  dimnames(Sex) <- list(Sex = c("F", "M"),
                    Group = c("ET_B","ET_P"))
    
  p1 <- chisq.test(Sex)$p.value
  
  # Age
  p2 <- wilcox.test(dat.cas$Age, dat.con$Age)$p.value

  # Weight
  p3 <- wilcox.test(dat.cas$Weight, dat.con$Weight)$p.value
  
  # BMI
  p4 <- wilcox.test(dat.cas$BMI, dat.con$BMI)$p.value

  p <- data.frame(
             "Number" = nrow(dat),
             "Sex" = p1, 
             "Age" = p2, 
             "Weight" = p3, 
             "BMI" = p4) %>% 
    t() %>% data.frame() %>%
    setNames("cohort\nWilcox.test Chisq.test")
}

# loss
lossFun <- function(x, tag){
 dat <- x %>% filter(Stage%in%tag) %>% 
        select(-one_of(c("ID","SampleID","Sex","Stage","Age")))  
 name <- c(paste0(tag, grp, "\nMean+/-SD"), paste0(tag, " pvalue"))
 # mean +/- sd ; p vlaue 
  res <- apply(dat %>% select(-Group), 2, function(x){
    # mean sd
    md <- tapply(x, dat$Group, function(y){
      mn <- paste( round(mean(y), 2), round(sd(y), 2), sep="+/-")
      return(mn)
      })
    # p value 
    pv <- wilcox.test(x~dat$Group)$p.value
    res <- c(md, pv)
    }) %>% data.frame() %>% t() %>% data.frame() %>% 
    setNames(name) 
  
  return(res)
}
```

### cohort description 
```{r}
cohortin <- CohortInform(phen)
DT::datatable(cohortin)
```

### Effect of CR on BMI (kg/m^2)

1. loss ratio  (before - after)*100 %
```{r, fig.width=4, fig.height=5}
# Weight/BMI loss   
loss <- phe.grp %>% select(ID, Stage, BMI, Weight) %>% 
  arrange(ID, Stage) %>% 
  group_by(ID) %>% 
  summarise(weight.k=(first(Weight)-last(Weight)),
            weight.p=round((first(Weight)-last(Weight))/first(Weight), 4)*100,
            BMI.k=(first(BMI)-last(BMI)),
            BMI.p=round((first(BMI)-last(BMI))/first(BMI), 4)*100) %>%
  ungroup() %>%
  arrange(BMI.k, ID)

phe.loss <- inner_join(phe.grp, loss, by="ID") %>% 
       arrange(ID, Stage) 
```


### Table 1 Cohort Description in two Enterotypes

 * Mean +/- SD 

```{r}
cohort.p <- CohortP(phe.loss) 
F.des <- CohortNum(phe.loss, grp[1])
M.des <- CohortNum(phe.loss, grp[2])
cohort.des <- cbind(F.des, M.des)
cohort <- cbind(cohort.des, cohort.p)
DT::datatable(cohort)
```

### Table 2 Effect of Calorie Restriction (CR) on BMI (kg/m^2)
```{r}
before <- phe.loss %>% filter(Stage%in%stg[1])
before.e <- lossFun(phe.loss, stg[1])
after.e <- lossFun(phe.loss, stg[2])
effect <- cbind(before.e, after.e)
DT::datatable(effect)
```

### output phentotype with group information
```{r}
write.csv(phe.loss, "../datset/phenotype/phenotype.grp.csv", row.names = F)
```
